---
title: "SageMath in R Markdown: One More Love Story"
author: | 
  | Yuanting Lu
  | Mercer University
date: "May 28, 2020"
output: html_document
---
<style type="text/css">

body{ /* Normal  */
      font-size: 20px;
      font-family: "Corbel";
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
    font-family: "Candara";
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

1. Create a blank sageHeader.html file, then paste the following HTML codes and save the file. The codes defines two HTML classes:
* The class "compute" has a Sage code editor, an evaluation button and the output panel.
* The class "simple" only displays the output panel.
The codes also send all calculations to [**SageMathCells** server](https://sagecell.sagemath.org/). Thus, the local installation of Sage is not needed.
```{r eval=FALSE, echo=TRUE}
<!DOCTYPE HTML>
<html>
  <head>
    <meta charset="utf-8">
    <meta name="viewport" content="width=device-width">
    <script src="https://sagecell.sagemath.org/static/embedded_sagecell.js"></script>
    <script>$(function () {
    sagecell.makeSagecell({inputLocation: 'div.compute',
                                  linked: 'false',
                                autoeval: 'false',
                          evalButtonText: 'Evaluate'});
    sagecell.makeSagecell({inputLocation:  'div.simple',
			                          autoeval: 'true',
			                              hide: ['editor','evalButton', 'permalink']});
    });
    </script>
  </head>
</html>
```

2. Create a R Markdown file and Source the sageHeader.html file in a R chunk using the **htmltools** package. Suppose the R Markdown file and the html file are in the same folder:
```{r eval=TRUE, echo=FALSE}
htmltools::includeHTML("sageHeader.html")
```
````markdown
`r ''````{r eval=TRUE, echo=FALSE}
htmltools::includeHTML("~/Desktop/sageHeader.html")
```
````

3. If we'd like the audience to type/modify the Sage codes, use the "**compute**" class. Insert the following HTML codes in the R Markdown document (don't use any code chunk) to get a Sage editor and an evaluation buttonin. (There is a button in the far right of the editor, which makes the editor full screen for easy editing.)
```{r eval = FALSE, echo=TRUE}
<html>
<div class="compute"><script type="text/x-sage">
factorial(5)
</script></div>
</html>
```
<html>
<div class="compute"><script type="text/x-sage">
factorial(5)
</script></div>
</html>

4. If we only want to display a dynamic output (e.g., a 3D graph),, use the "**simple**" class. The following HTML codes in the R Markdown document (don't use any code chunk) will only display the result of the Sage codes.  (It may take a while for the image to show up. Refresh the browser might help.)
```{r eval=FALSE, echo=TRUE}
<html>
<div class="simple"><script type="text/x-sage">
# Declare variables
var('u,v') 
# Choose the jet colormap. 
color_map = colormaps.jet 
# c is the color function, whose range must be normalized (converted to between 0 and 1). 
def c(u,v): return ((u^2-v^2)+1)/2 
# 3d plot and settings
surf = parametric_plot3d( [u, v, u^2-v^2], (u, -1, 1), (v, -1, 1), color = (c, color_map)) 
show(surf, viewer='threejs', frame=False) 
</script></div>
</html>
```
<html>
<div class="simple"><script type="text/x-sage">
var('u,v'); color_map = colormaps.jet          
def c(u,v): return ((u^2-v^2)+1)/2 
surf = parametric_plot3d( [u, v, u^2-v^2], (u, -1, 1), (v, -1, 1), color = (c, color_map))
show(surf, viewer='threejs', frame=False)
</script></div>
</html>

5. Finally, don't forget the wonderful interactive components of Sage. 
```{r eval=TRUE, echo=FALSE}
htmltools::includeHTML("riemannsum.html")
```


# References
* [SageMath and 3D Models in Webpages, by Simon Willerton](https://golem.ph.utexas.edu/category/2017/12/sagemath_and_3d_models_in_webp.html)
* [SageMath Widgets, by Jeff Denny and Yuanting Lu](http://sagewidgets.pbworks.com/w/page/67137878/FrontPage)
* [Dr. Christian Bar's colletion, University of Potsdam, Germany](https://www.cbaer.eu/joomla/index.php/en/mathematics/practice)
* [Embedding Sage to HTML, by SageMath](https://sagecell.sagemath.org/static/about.html?v=aae960fc94126c35e9d68d7f5d2040a7)
* [R Markdown font size, answered by Mike Wise on StackOverFlow ](https://stackoverflow.com/questions/30446905/rmarkdown-font-size-and-header/30447045)
* [Including verbatim R code chunks inside R Markdown, by Nathan Stephens](https://support.rstudio.com/hc/en-us/articles/360018181633-Including-verbatim-R-code-chunks-inside-R-Markdown)